---
title: "Appendix 1 - King County house price prediction"
output:
  html_notebook: default
  word_document: default
---

## Part 1 - Multi-Linear Regression Model

### Step 1 : Collecting Data
The data set is collected in CSV format from kaggele and below is the reference for the same.

Harlfoxem. (2016) House sales in king county, usa. [Online]. Available:https://www.kaggle.com/harlfoxem/housesalesprediction

### Step 2 : Exploring, preprocessing and cleaning the data
Primary setup
```{r setup}
knitr::opts_knit$set(root.dir = '/Users/sobil/Documents/MSC/Sem 1/Data Mining & Machine Learning/Project/Kc_house/')
remove(list = ls())
```

#### 1) reading the raw csv file
```{r}
kc_house <- read.csv("kc_house_data.csv", stringsAsFactors = FALSE)
```

#### 2) exploratory analysis 
Structure of the kc_house data frame
```{r}
str(kc_house)
```

Summary of the kc_house data frame
```{r}
summary(kc_house)
```
Plotting histogram of the house prices
```{r}
library(ggplot2)
library(ggthemes)
theme_set(theme_gdocs())
ggplot(kc_house, aes(x=price)) + geom_histogram(bins = 10, color = 'blue', aes(fill=..count..), alpha = 0.4) + xlab('House Price') + ylab('Count') + ggtitle('House Price distribution Plot')
```

#### 3) cleaning & removing the unrelated and insignificant columns
removing id
```{r}
kc_house <- kc_house[,-c(1)]
```

removing the lat, long, zipcode
```{r}
kc_house <- kc_house[,-c(18,16,17)]
```

sqft_living, sqft_lot are related and can calculate sqft_living15, sqft_lot15 resp.
```{r}
kc_house <- kc_house[,-c(17,16)]
```

cheking NAs
```{r}
apply(X = kc_house,MARGIN =  2, FUN = function(col) any(is.na(col))) # no NAs
```

#### 4) exploring the colmuns for outliers and influential points
```{r}
table(kc_house$bedrooms) 
```
33 bedroom seems to be a influential point
```{r}
kc_house <- kc_house[kc_house$bedrooms != 33,]
```

Column - bathrooms
```{r}
table(kc_house$bathrooms)
```

Column - View
```{r}
table(kc_house$view)
```

Resetting the row names to index
```{r}
row.names(kc_house) <- 1:21612
```

### Step 3.1 - Data transformation - default price

coverting the date field type from chr to date
```{r}
kc_house$date <- as.Date(kc_house$date, format = "%Y%m%d")
summary(kc_house$date)
```

Date is almost for the same duration of time, so the sample is unbiased
remove date
```{r}
kc_house <- kc_house[,-c(1)]
```

converting the waterfront to factors (0,1)
```{r}
table(kc_house$waterfront)
kc_house$waterfront <- as.factor(kc_house$waterfront)
```

### Step 4.1 : Training a model on the data - default price

adding the libraries required
```{r}
library(boot)
library(car)
library(leaps)
library(psych)
```

checking correlation
```{r}
pairs.panels(kc_house)
```

Model the linear model
simple model 1
```{r}
house.fit1 <- glm(price ~ ., data = kc_house)
```

### Step 5.1 : Evaluating the model
Checking the summary of the model
```{r}
summary(house.fit1)
```

### Step 6.1 : Improving the model
#### 1) Adding all the possible relation to the model
Model 2 - step backward
```{r}
house.fit2 <- step(house.fit1)
summary(house.fit2)
```

Model 3 - adding interaction
```{r}
house.fit3 <- update(house.fit2, ~ . + bedrooms:bathrooms + bedrooms:grade + bedrooms:sqft_above + bedrooms:sqft_living + bathrooms:grade + bathrooms:sqft_above + sqft_above:sqft_living + grade:sqft_living + bathrooms:sqft_living + floors:sqft_living + view:sqft_living + floors:condition + floors:grade + floors:sqft_above + waterfront:view + view:grade)
summary(house.fit3)
```

Model 4 - step backward
```{r}
house.fit4 <- step(house.fit3)
summary(house.fit4)
```

Model 5 - removing insigificant floors predictor
```{r}
house.fit5 <- update(house.fit2, ~ . - floors)
summary(house.fit5)
```

#### 2) check the best fit by ignoring uncorrelated variables
checking the best predictors to predict the house price
```{r}
bstFits1 <- regsubsets(price ~ bedrooms+bathrooms+sqft_living+sqft_lot+floors+waterfront+view+condition+grade+sqft_above+sqft_basement+yr_built+yr_renovated, data = kc_house, nbest = 1, nvmax = 6)
par(mfrow = c(1,1))
#subsets(bstFits1, statistic = "adjr2", max.size = 6)
plot(bstFits1, scale = "adjr2")
```

Doing cross-validation for 4 predictors using K-Fold
```{r}
house.k10.4 <- glm(price ~ sqft_living + waterfront + yr_renovated + grade, data = kc_house)
summary(house.k10.4) # 596532
house.k10.4.err <- cv.glm(data = kc_house,glmfit = house.k10.4, K = 10)
round(house.k10.4.err$delta[1], 4) # 56952103727
```

Checking the r square and adj. r square values
```{r}
house.k10.4.lm <- lm(price ~ sqft_living + waterfront + yr_renovated + grade, data = kc_house)
summary(house.k10.4.lm)
```

Plot the model
```{r}
par(mfrow = c(2,2))
plot(house.k10.4)
```
### Step 3.2 - Data transformation - log(price) and using all variables
#### After looking at the results of the plot and previously from histogram in step 2 - taking log of the price

checking distribution & outliers of dependent value
```{r}
par(mfrow = c(1,2))
hist(kc_house$price,
     col="orange", 
     border="black",
     prob = TRUE,
     xlab = "House prices",
     main = "Histogram")
lines(density(kc_house$price),
      lwd = 2,
      col = "chocolate3") # not distributed normally - right skewed
plot(kc_house$price, main = "Scatter plot", ylab = "House prices")

# covert to log
hist(log(kc_house$price),
     col="orange", 
     border="black",
     prob = TRUE,
     xlab = "House prices",
     main = "Histogram")
lines(density(log(kc_house$price)),
      lwd = 2,
      col = "chocolate3") # not distributed normally
plot(log(kc_house$price), main = "Scatter plot", ylab = "House prices")
```

creating a new data frame - coverting price to log(price)
```{r}
kc_house_log <- kc_house 
kc_house_log$price <- log(kc_house_log$price)
summary(kc_house_log)
```

### Step 4.2 : Training a model on the data - log(price) and using all variables

checking correlation
```{r}
pairs.panels(kc_house_log)
```

model 1
```{r}
house.glm1 <- glm(price ~ bedrooms+bathrooms+sqft_living+sqft_lot+floors+waterfront+view+condition+grade+sqft_above+sqft_basement+yr_built+yr_renovated, data = kc_house_log)
```

### Step 5.2 : Evaluating the model - log(price) and using all variables
```{r}
summary(house.glm1)
```

### Step 6.2 : Improving the model - log(price) and using all variables
#### 1) Adding all the possible relation to the model
model 2
```{r}
house.glm2 <- update(house.glm1, ~ . - sqft_basement)
summary(house.glm2)
```

model 3
```{r}
house.glm3 <- update(house.glm2, ~ . - sqft_lot)
summary(house.glm3)
```

model 4
```{r}
house.glm4 <- update(house.glm3, ~ . - yr_renovated)
summary(house.glm4)
```

adding interaction by checking the graph
model 5 - sqft_living interaction
```{r}
house.glm5 <- update(house.glm4, ~ . + sqft_above:sqft_living + grade:sqft_living + bathrooms:sqft_living + floors:sqft_living + view:sqft_living)
summary(house.glm5)
```

model 6
```{r}
house.glm6 <- update(house.glm5, ~ . - bathrooms:sqft_living)
summary(house.glm6)
```

model 7 - bedrooms interaction
```{r}
house.glm7 <- update(house.glm6, ~ . + bedrooms:bathrooms + bedrooms:grade + bedrooms:sqft_above)
summary(house.glm7)
```

 model 8
```{r}
house.glm8 <- update(house.glm7, ~ . - bedrooms:sqft_above - bedrooms:bathrooms - sqft_living:sqft_above)
summary(house.glm8)
```

model 9 - bathrooms interaction
```{r}
house.glm9 <- update(house.glm8, ~ . + bathrooms:floors + bathrooms:grade + bathrooms:sqft_above)
summary(house.glm9)
```

model 10
```{r}
house.glm10 <- update(house.glm9, ~ . - bathrooms:grade - bathrooms:floors)
summary(house.glm10)
```

model 11 - floors interaction
```{r}
house.glm11 <- update(house.glm10, ~ . + floors:condition + floors:grade + floors:sqft_above)
summary(house.glm11)
```

model 12
```{r}
house.glm12 <- update(house.glm11, ~ . - sqft_living:floors - floors:condition -floors:grade)
summary(house.glm12)
```

model 13
```{r}
house.glm13 <- update(house.glm12, ~ . - sqft_above)
summary(house.glm13)
```

model 14 - floors interaction
```{r}
house.glm14 <- update(house.glm13, ~ . + waterfront:view + view:grade + grade:sqft_above)
summary(house.glm14)
```

model 15
```{r}
house.glm15 <- update(house.glm14, ~ . - grade:sqft_above - view:grade - waterfront)
summary(house.glm15)
```


#### 2) as per concept of parsimony : checking best fit from above model
```{r}
house.glm15.bestFit1 <- regsubsets(price ~ bedrooms + bathrooms + sqft_living + floors + 
                                  view + condition + grade + yr_built + sqft_living:grade + 
                                  sqft_living:view + bedrooms:grade + bathrooms:sqft_above + 
                                  floors:sqft_above + waterfront:view, data = kc_house_log, nbest = 2, nvmax = 10)
par(mfrow = c(1,1))
#subsets(house.glm15.bestFit1, statistic = "adjr2")
plot(house.glm15.bestFit1, scale = "adjr2")
```

from above we can see 4,5,6 and later produce almost similar result
```{r}
house.glm15.bestFit2 <- regsubsets(price ~ bedrooms + bathrooms + sqft_living + floors + 
                                     view + condition + grade + yr_built + sqft_living:grade + 
                                     sqft_living:view + bedrooms:grade + bathrooms:sqft_above + 
                                     floors:sqft_above + waterfront:view, data = kc_house_log, nbest = 1, nvmax = 6)
par(mfrow = c(1,1))
#subsets(house.glm15.bestFit2, statistic = "adjr2", max.size = 6, min.size = 1)
plot(house.glm15.bestFit2, scale = "adjr2")
```

Doing cross-validation for 3,4,5,6 predictors using K-Fold
3 predictors
predictor : 3-i
```{r}
house.k10.3i <- glm(price ~ sqft_living + grade + yr_built, data = kc_house_log)
summary(house.k10.3i) # 12786
house.k10.3i.err <- cv.glm(data = kc_house_log,glmfit = house.k10.3i, K = 10)
round(house.k10.3i.err$delta[1], 4) # 0.1058
```

predictor : 3-ii
```{r}
house.k10.3ii <- glm(price ~ sqft_living:grade + grade + yr_built, data = kc_house_log)
summary(house.k10.3ii) # 13539
house.k10.3ii.err <- cv.glm(data = kc_house_log,glmfit = house.k10.3ii, K = 10)
round(house.k10.3ii.err$delta[1], 4) # 0.1096
```

4 predictors
predictor : 4-i
```{r}
house.k10.4i <- glm(price ~ sqft_living + grade + yr_built + view, data = kc_house_log)
summary(house.k10.4i) # 12272
house.k10.4i.err <- cv.glm(data = kc_house_log,glmfit = house.k10.4i, K = 10)
round(house.k10.4i.err$delta[1], 4) # 0.1033
```

predictor : 4-ii
```{r}
house.k10.4ii <- glm(price ~ sqft_living + grade + yr_built + bathrooms, data = kc_house_log)
summary(house.k10.4ii) # 12351
house.k10.4ii.err <- cv.glm(data = kc_house_log,glmfit = house.k10.4ii, K = 10)
round(house.k10.4ii.err$delta[1], 4) # 0.1037
```

5 predictors
predictor : 5-i
```{r}
house.k10.5i <- glm(price ~ sqft_living + grade + yr_built + view + bathrooms, data = kc_house_log)
summary(house.k10.5i) # 11844
house.k10.5i.err <- cv.glm(data = kc_house_log,glmfit = house.k10.5i, K = 10)
round(house.k10.5i.err$delta[1], 4) # 0.1013
```

predictor : 5-ii
```{r}
house.k10.5ii <- glm(price ~ sqft_living + grade + yr_built + view + bathrooms, data = kc_house_log)
summary(house.k10.5ii) # 11844
house.k10.5ii.err <- cv.glm(data = kc_house_log,glmfit = house.k10.5ii, K = 10)
round(house.k10.5ii.err$delta[1], 4) # 0.1013
```

6 predictors
predictor : 6-i
```{r}
house.k10.6i <- glm(price ~ sqft_living + grade + yr_built + view + bathrooms + floors, data = kc_house_log)
summary(house.k10.6i) # 11587
house.k10.6i.err <- cv.glm(data = kc_house_log,glmfit = house.k10.6i, K = 10)
round(house.k10.6i.err$delta[1], 4) # 0.1001
```

predictor : 6-ii
```{r}
house.k10.6ii <- glm(price ~ sqft_living + grade + yr_built + view + sqft_living:grade + floors, data = kc_house_log)
summary(house.k10.6ii) # 11589
house.k10.6ii.err <- cv.glm(data = kc_house_log,glmfit = house.k10.6ii, K = 10)
round(house.k10.6ii.err$delta[1], 4) # 0.1001
```

Checking the r square and adj. r square values for house.k10.6i model
```{r}
house.k10.6i.lm <- lm(price ~ sqft_living + grade + yr_built + view + bathrooms + floors, data = kc_house_log)
summary(house.k10.6i.lm)
```

Plot the model for house.k10.6i.lm model
```{r}
par(mfrow = c(2,2))
plot(house.k10.4)
```

### Step 3.3 - Data transformation - log(price) conversion and using only numeric type parameters
removing very less correlated variable by creating a new df
```{r}
house <- kc_house_log
```
remove sqft_lot, floors, waterfront, view, condition, sqft_basement, yr_built, yr_renovated i.e columns (5,6,7,8,9,13,14)

```{r}
house <- house[,-c(5,6,7,8,9,12,13,14)]
```

### Step 4.3 : Training a model on the data - log(price) conversion and using only numeric type parameters

Model the linear model
model fit 1
```{r}
house.fit1 <- lm(price ~ bedrooms*bathrooms*sqft_living*grade*sqft_above, data = house)
```

### Step 5.3 : Evaluating the model - log(price) conversion and using only numeric type parameters
Checking the summary and plot of the model
```{r}
summary(house.fit1)
par(mfrow = c(2,2))
plot(house.fit1)
```

### Step 6.3 : Improving the model - log(price) conversion and using only numeric type parameters
#### 1) Adding all the possible relation to the model
fit 2
```{r}
house.fit2 <- update(house.fit1, ~ . + I(bedrooms^2) + I(bathrooms^2) + I(sqft_living^2) + I(grade^2) + I(sqft_above^2))
summary(house.fit2)
```

fit 3
```{r}
house.fit3 <- update(house.fit2, ~ . + I(bedrooms^3) + I(bathrooms^3) + I(sqft_living^3) + I(grade^3) + I(sqft_above^3))
summary(house.fit3)
```

fit 4
```{r}
house.fit4 <- update(house.fit3, ~ . + I(bedrooms^0.5) + I(bathrooms^0.5) + I(sqft_living^0.5) + I(grade^0.5) + I(sqft_above^0.5))
summary(house.fit4)
```

fit 4 : step function
```{r}
house.fit5 <- step(house.fit4)
summary(house.fit5)
```

fit 6 :  removing non - significant coefficients manually
```{r}
house.fit6 <- update(house.fit5, ~ . - bedrooms:sqft_above - bedrooms:sqft_living - bedrooms:bathrooms)
summary(house.fit6)
```

fit 7 : step function
```{r}
house.fit7 <- update(house.fit6, ~ . - bathrooms)
summary(house.fit7)
```

#### 2) check the best fit by ignoring uncorrelated variables
from model 7 formula : 
lm(formula = price ~ bedrooms + sqft_living + grade + sqft_above + 
 I(bedrooms^2) + I(bathrooms^2) + I(sqft_living^2) + I(sqft_above^2) + 
    I(bedrooms^3) + I(bathrooms^3) + I(sqft_living^3) + I(grade^3) + 
    I(sqft_above^3) + I(bedrooms^0.5) + I(bathrooms^0.5) + I(grade^0.5) + 
    bathrooms:sqft_living + bedrooms:grade + bathrooms:grade + 
    sqft_living:grade + bathrooms:sqft_above + sqft_living:sqft_above + 
    grade:sqft_above + bedrooms:bathrooms:sqft_living + bedrooms:bathrooms:sqft_above + 
    bathrooms:sqft_living:sqft_above + bathrooms:grade:sqft_above + 
    sqft_living:grade:sqft_above, data = house) 


check the best fit
```{r}
bstFits1 <- regsubsets(price ~ bedrooms + sqft_living + grade + sqft_above + 
                         I(bedrooms^2) + I(bathrooms^2) + I(sqft_living^2) + I(sqft_above^2) + 
                         I(bedrooms^3) + I(bathrooms^3) + I(sqft_living^3) + I(grade^3) + 
                         I(sqft_above^3) + I(bedrooms^0.5) + I(bathrooms^0.5) + I(grade^0.5) + 
                         bathrooms:sqft_living + bedrooms:grade + bathrooms:grade + 
                         sqft_living:grade + bathrooms:sqft_above + sqft_living:sqft_above + 
                         grade:sqft_above + bedrooms:bathrooms:sqft_living + bedrooms:bathrooms:sqft_above + 
                         bathrooms:sqft_living:sqft_above + bathrooms:grade:sqft_above + 
                         sqft_living:grade:sqft_above, data = house, nbest = 1, nvmax = 4)
par(mfrow = c(1,1))
#subsets(bstFits1, statistic = "adjr2")
plot(bstFits1, scale = "adjr2")
```

Final model - 
```{r}
house.final <- lm (price ~ sqft_living + grade + sqft_above, data = house)
summary(house.final)
par(mfrow = c(2,2))
plot(house.final)
```

removing the influencial points
```{r}
house <- house[- c(12778),]
```

same model again
```{r}
house.final <- lm (price ~ sqft_living + grade + sqft_above, data = house)
summary(house.final)
par(mfrow = c(2,2))
plot(house.final)
```

checking the accuracy using k-Fold method
```{r}
library(boot)
house.glm <- glm(price ~ sqft_living + grade + sqft_above, data = house)
summary(house.glm) # AIC = 15670
house.glm.err<- cv.glm(data = house,glmfit = house.glm, K = 10)
round(house.glm.err$delta[1],4)# 0.1209
```
