---
title: "Appendix 3 - Customer’s  decision  to  open  a  term  deposit account"
output:
  word_document: default
  html_notebook: default
---

## Part 3 - Decision Tree

### Step 1 : Collecting Data
Same as of Part 1

### Step 2 : Exploring, preprocessing and cleaning the data
Primary setup
```{r setup}
knitr::opts_knit$set(root.dir = '/Users/sobil/Documents/MSC/Sem 1/Data Mining & Machine Learning/Project/Bank/')
remove(list = ls())
set.seed(1)
```

loading all the libraries required
```{r}
library(caret)
library(C50)
library(RWeka)
```

rest same as part 1
```{r}
source("bank_import_primaryExplore.R")
```

### Step 3 - Data transformation & preparation
creating training and testing dataset from exisitng sample
```{r}
set.seed(1)
indx <- createDataPartition(bank$y, p = 0.8, list = FALSE)

bank_train <- bank[indx,]
bank_test <- bank[- indx,]

prop.table(table(bank_train$y))
prop.table(table(bank_test$y))
```

### Step 4 : Train the model
```{r}
bank_model <- C50::C5.0(bank_train[-21],bank_train$y)
```

### Step 5 : Evaluate the model
```{r}
bank_model
bank_predict <- predict(object = bank_model, newdata = bank_test)
caret::confusionMatrix(bank_predict, bank_test$y)
```

### Step 6 : Improving model performance
#### 1) Method 1 - Boosting trials = 10
```{r}
bank_boost10 <- C50::C5.0(bank_train[-21], bank_train$y, trials = 10)
bank_boost10

bank_boost_predict10 <- predict(object = bank_boost10, newdata = bank_test)
caret::confusionMatrix(bank_boost_predict10, bank_test$y)
```

#### 2) Method 2 - Boosting trials = 5
```{r}
bank_boost5 <- C50::C5.0(bank_train[-21], bank_train$y, trials = 5)
bank_boost5

bank_boost_predict5 <- predict(object = bank_boost5, newdata = bank_test)
caret::confusionMatrix(bank_boost_predict5, bank_test$y)
```

#### 3) Method 3 - Adding Cost matrix
creating cost matrix - "yes" more costlier than others
```{r}
mtr_dim <- list(c("no", "yes"), c("no","yes"))
names(mtr_dim) <- c("predict","actual")
mtr_dim

err_cst <- matrix(c(0,1,0,4), nrow = 2, dimnames = mtr_dim)
err_cst
```

Model Costs
```{r}
bank_cost <- C50::C5.0(bank_train[-21], bank_train$y, costs = err_cst)
bank_cost_predict <- predict(object = bank_cost, newdata = bank_test)
caret::confusionMatrix(bank_cost_predict, bank_test$y)
```

#### 4) Method 4 - 1R algo model
```{r}
bank_1R <- OneR(y ~ ., data = bank_train)
bank_1R
bank_predict_1R <- predict(object = bank_1R, newdata = bank_test)
caret::confusionMatrix(bank_predict_1R, bank_test$y)
```

#### 5) Method 5 - RIPPER algo model
```{r}
bank_JRip <- JRip(y ~ . , data = bank_train)
bank_JRip
bank_predict_JRip <- predict(object = bank_JRip, newdata = bank_test)
caret::confusionMatrix(bank_predict_JRip, bank_test$y)
#kappa = 0.5667,   Sensitivity = 0.58513,    Specificity = 0.95895 ====> BEST
```
```{r}
caret::confusionMatrix(bank_predict_JRip, bank_test$y, mode="prec_recall")
```

