---
title: "Appendix 3 - Customer’s  decision  to  open  a  term  deposit account"
output:
  html_notebook: default
  word_document: default
---

## Part 1 - KNN Model

### Step 1 : Collecting Data
The data set is collected in CSV format from UCI.edu and below is the reference for the same.

S. Moro, P. Cortez, and P. Rita. (2014) A data-driven approach to predictthe success of bank telemarketing. Decision Support Systems,. [Online].Available: https://archive.ics.uci.edu/ml/datasets/bank marketing

### Step 2 : Exploring, preprocessing and cleaning the data
Primary setup
```{r setup}
knitr::opts_knit$set(root.dir = '/Users/sobil/Documents/MSC/Sem 1/Data Mining & Machine Learning/Project/Bank/')
remove(list = ls())
set.seed(1)
```

loading all the libraries required
```{r}
library(gmodels)
library(caret)
library(dplyr)
library(class)
library(fastDummies)
```

#### 1) reading the raw csv file
```{r}
bank <- read.csv("bank-additional-full.csv", sep = ";")
```

#### 2) exploratory analysis 
Structure of the bank data frame
```{r}
str(bank)
```

Summary of the bank data frame
```{r}
summary(bank)
```

checking the overall distribution of prediction/ target charcterstics/feature
```{r}
gmodels::CrossTable(bank$y)
```

converting the target feature into factor with adding labels for more information
```{r}
bank$y <- factor(bank$y, levels = c("yes", "no"), labels = c("Yes", "No"))
```

cheking NAs
```{r}
apply(X = bank,MARGIN =  2, FUN = function(col) any(is.na(col))) # no NAs
```

### Step 3.1 - Data transformation & preparation - Normalizing the data with custom function

normalizing the data to standardize range of values of various characterstics
#### 1) creating normalized method
```{r}
nor <- function(val) {
  return((val - min(val))/(max(val) - min(val)))
}
```

getting normalised variables in a data frame
```{r}
bank_n_cont <- as.data.frame(lapply(bank[c(1,11,12,13,14,16,17,18,19,20)], nor))
summary(bank_n_cont)
```

#### 2) creating a new df of only independent variables with normalised columns
```{r}
bank_n <- bank[,-21]
bank_n$age <- bank_n_cont$age
bank_n$duration <- bank_n_cont$duration
bank_n$campaign <- bank_n_cont$campaign
bank_n$pdays <- bank_n_cont$pdays
bank_n$previous <- bank_n_cont$previous
bank_n$emp.var.rate <- bank_n_cont$emp.var.rate
bank_n$cons.price.idx <- bank_n_cont$cons.price.idx
bank_n$cons.conf.idx <- bank_n_cont$cons.conf.idx
bank_n$euribor3m <- bank_n_cont$euribor3m
bank_n$nr.employed <- bank_n_cont$nr.employed
summary(bank_n)
```

#### 3) creating dummy columns for categorical columns
```{r}
bank_n <- fastDummies::dummy_cols(bank_n, remove_first_dummy = TRUE) # remove_first_dummy = TRUE to avoid multi-collinearity
```

removing categorical columns from df (dummy are included)
```{r}
bank_n <- bank_n[,-c(2:10,15)]
```

#### 4) creating training and testing dataset from exisitng sample
```{r}
indx <- createDataPartition(bank$y, p = 0.8, list = FALSE)
bank_train <- bank_n[indx,]
bank_test <- bank_n[- indx,]
```

creating lables for test and training data sets
```{r}
bank_train_labels <- bank[indx,21]
bank_test_labels <- bank[- indx,21]
```

### Step 4.1 : Training a model on the data - Normalizing the data with custom function
using the knn method of class package with K value equalent to square root of total train observation & odd number to eliminate tie vote issue for 2 factor classification i.e. 
```{r}
bank$y %>% length %>% sqrt %>% round # 203
bank_test_predict <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 203)
```

### Step 5.1 : Evaluating the model - Normalizing the data with custom function
"Model 1 : k = 203"
```{r}
confusionMatrix(data = bank_test_predict,reference = bank_test_labels)
```

### Step 6.1 : Improving the model
#### 1) Changing k values

```{r}
bank_test_predict2 <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 151)
print("Model 2 : k = 151")
print(confusionMatrix(data = bank_test_predict2,reference = bank_test_labels))
```

predict = 3     : k = 101
```{r}
bank_test_predict3 <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 101)
print("Model 3 : k = 101")
print(confusionMatrix(data = bank_test_predict3,reference = bank_test_labels))

bank_test_predict4 <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 75)
print("Model 4 : k = 75")
print(confusionMatrix(data = bank_test_predict4,reference = bank_test_labels))

bank_test_predict5 <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 51)
print("Model 5 : k = 51")
print(confusionMatrix(data = bank_test_predict5,reference = bank_test_labels))

bank_test_predict6 <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 31)
print("Model 6 : k = 31")
print(confusionMatrix(data = bank_test_predict6,reference = bank_test_labels))

bank_test_predict7 <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 21)
print("Model 7 : k = 21")
print(confusionMatrix(data = bank_test_predict7,reference = bank_test_labels))

bank_test_predict8 <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 11)
print("Model 8 : k = 11")
print(confusionMatrix(data = bank_test_predict8,reference = bank_test_labels))
```

Model n : k = 9 to 1 (only odd numbers)
```{r}
i <- 9
while (i > 0) {
  bank_test_predict_n <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = i)
  print(i)
  print(confusionMatrix(data = bank_test_predict_n,reference = bank_test_labels))
  i <- i - 2
}
```

Model n : k = 19 to 13 (only odd numbers)
```{r}
i <- 19
while (i > 11) {
  bank_test_predict_n <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = i)
  print(i)
  print(confusionMatrix(data = bank_test_predict_n,reference = bank_test_labels))
  i <- i - 2
}
```

Model n : k = 29 to 23 (only odd numbers)
```{r}
i <- 29
while (i > 21) {
  bank_test_predict_n <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = i)
  print(i)
  print(confusionMatrix(data = bank_test_predict_n,reference = bank_test_labels))
  i <- i - 2
}
```

#### 2) Best among all is for K = 3
```{r}
bank_test_predict.best <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 3)
print(confusionMatrix(data = bank_test_predict.best,reference = bank_test_labels))
```

### Step 3.2 - Data transformation & preparation - Normalizing the data with inbuilt R function

Redoing the first 2 steps before transformation
```{r}
source("/Users/sobil/Documents/MSC/Sem 1/Data Mining & Machine Learning/Project/Bank/bank_import_primaryExplore.R")
```

normalizing the data to standardize range of values of various characterstics

getting normalised variables in a data frame

#### 1) Normalization using scale inbulit R method based upon using z score
```{r}
bank_n_cont <- as.data.frame(scale(bank[c(1,11,12,13,14,16,17,18,19,20)]))
summary(bank_n_cont)
```

#### 2) creating a new df of only independent variables
```{r}
bank_n <- bank[,-21]
bank_n$age <- bank_n_cont$age
bank_n$duration <- bank_n_cont$duration
bank_n$campaign <- bank_n_cont$campaign
bank_n$pdays <- bank_n_cont$pdays
bank_n$previous <- bank_n_cont$previous
bank_n$emp.var.rate <- bank_n_cont$emp.var.rate
bank_n$cons.price.idx <- bank_n_cont$cons.price.idx
bank_n$cons.conf.idx <- bank_n_cont$cons.conf.idx
bank_n$euribor3m <- bank_n_cont$euribor3m
bank_n$nr.employed <- bank_n_cont$nr.employed
summary(bank_n)
```

#### 3) creating dummy columns for categorical columns
```{r}
bank_n <- fastDummies::dummy_cols(bank_n, remove_first_dummy = TRUE) # remove_first_dummy = TRUE to avoid multi-collinearity
```

removing categorical columns from df (dummy are included)
```{r}
bank_n <- bank_n[,-c(2:10,15)]
```


#### 4) creating training and testing dataset from exisitng sample
```{r}
indx <- createDataPartition(bank$y, p = 0.8, list = FALSE)
bank_train <- bank_n[indx,]
bank_test <- bank_n[- indx,]
```

creating lables for test and training data sets
```{r}
bank_train_labels <- bank[indx,21]
bank_test_labels <- bank[- indx,21]
```

### Step 4.2 : Training a model on the data - Normalizing the data with inbuilt R function
using the knn method of class package with K odd 
```{r}
bank_test_predict <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 51)
```

### Step 5.2 : Evaluating the model - Normalizing the data with inbuilt R function
"Model 1 : k = 51"
```{r}
confusionMatrix(data = bank_test_predict,reference = bank_test_labels)
```

### Step 6.2 : Improving the model - Normalizing the data with inbuilt R function
#### 1) Changing k values
Skipped model for K values which were higher than 51 (please refer the R script for model performance) - overall the performance was poor than below

Model n : k = 31 to 1 (only odd numbers)"
```{r}
i <- 31
while (i > 0) {
  bank_test_predict_n <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = i)
  print(i)
  print(confusionMatrix(data = bank_test_predict_n,reference = bank_test_labels))
  i <- i - 2
}
```

#### 2) Best among all is for K = 9
```{r}
bank_test_predict.best <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 9)
print(confusionMatrix(data = bank_test_predict.best,reference = bank_test_labels))
```

### Step 3.3 - Data transformation & preparation - Only numeric predictors

Redoing the first 2 steps before transformation
```{r}
source("/Users/sobil/Documents/MSC/Sem 1/Data Mining & Machine Learning/Project/Bank/bank_import_primaryExplore.R")
```

#### 1) Creating new data frame with only numeric predicors and normalizaing them
From previous section, it is clear that inbuilt normalization is performing better, Thus Normalization using scale inbulit R method based upon using z score
```{r}
bank_n <- as.data.frame(scale(bank[c(1,11,12,13,14,16,17,18,19,20)]))
summary(bank_n)
```


#### 2) creating training and testing dataset from exisitng sample
```{r}
indx <- createDataPartition(bank$y, p = 0.8, list = FALSE)
bank_train <- bank_n[indx,]
bank_test <- bank_n[- indx,]
```

creating lables for test and training data sets
```{r}
bank_train_labels <- bank[indx,21]
bank_test_labels <- bank[- indx,21]
```

### Step 4.3 : Training a model on the data - Only numeric predictors
using the knn method of class package with K odd 
```{r}
bank_test_predict <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 51)
```

### Step 5.3 : Evaluating the model - Only numeric predictors
"Model 1 : k = 51"
```{r}
confusionMatrix(data = bank_test_predict,reference = bank_test_labels)
```

### Step 6.3 : Improving the model - Only numeric predictors
#### 1) Changing k values
Skipped model for K values which were higher than 51 (please refer the R script for model performance) - overall the performance was poor than below

Model n : k = 31 to 1 (only odd numbers)"
```{r}
i <- 31
while (i > 0) {
  bank_test_predict_n <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = i)
  print(i)
  print(confusionMatrix(data = bank_test_predict_n,reference = bank_test_labels))
  i <- i - 2
}
```

#### 2) Best among all is for K = 19
```{r}
bank_test_predict.best <- class::knn(train = bank_train, test = bank_test, cl = bank_train_labels, k = 19)
print(confusionMatrix(data = bank_test_predict.best,reference = bank_test_labels))
```
```{r}
confusionMatrix(data = bank_test_predict.best,reference = bank_test_labels, mode="prec_recall")
```