---
title: "Appendix 3 - Customer’s  decision  to  open  a  term  deposit account"
output:
  html_notebook: default
  word_document: default
---

## Part 4 - SVM model

### Step 1 : Collecting Data
Same as of Part 1

### Step 2 : Exploring, preprocessing and cleaning the data
Primary setup
```{r setup}
knitr::opts_knit$set(root.dir = '/Users/sobil/Documents/MSC/Sem 1/Data Mining & Machine Learning/Project/Bank/')
remove(list = ls())
set.seed(1)
```

loading all the libraries required
```{r}
library(caret)
library(kernlab)
```

rest same as part 1
```{r}
source("bank_import_primaryExplore.R")
```

### Step 3 - Data transformation & preparation
creating training and testing dataset from exisitng sample
```{r}
set.seed(1)
indx <- createDataPartition(bank$y, p = 0.8, list = FALSE)

bank_train <- bank[indx,]
bank_test <- bank[- indx,]

prop.table(table(bank_train$y))
prop.table(table(bank_test$y))
```

### Step 4 : Training a model on the data
```{r}
bank_svm <- ksvm(y ~ ., data = bank_train, kernel = "vanilladot")
bank_svm
```

### Step 5 : Evaluating model performance
```{r}
bank_predict_svm <- predict(object = bank_svm, newdata = bank_test)
caret::confusionMatrix(bank_predict_svm, bank_test$y) # kappa = 0.4051,   Sensitivity = 0.33297,    Specificity = 0.98085
```

### Step 6 : Improving model performance
#### 1) Method 1 : kernel = rbfdot
```{r}
bank_svm.rbf <- ksvm(y ~ . , data = bank_train, kernel = "rbfdot")
bank_predict_svm.rbf <- predict(object = bank_svm.rbf, newdata = bank_test)
caret::confusionMatrix(bank_predict_svm.rbf, bank_test$y) # kappa = 0.4907,   Sensitivity = 0.42996,    Specificity = 0.97715 
```

#### Method 2 : kernel = polydot
```{r}
bank_svm.poly <- ksvm(y ~ . , data = bank_train, kernel = "polydot")
bank_predict_svm.poly <- predict(object = bank_svm.poly, newdata = bank_test)
caret::confusionMatrix(bank_predict_svm.poly, bank_test$y) # kappa = 0.4051,   Sensitivity = 0.33297,    Specificity = 0.98085 
```

#### Method 3 : kernel = tanhdot
```{r}
bank_svm.tanh <- ksvm(y ~ . , data = bank_train, kernel = "tanhdot")
bank_predict_svm.tanh <- predict(object = bank_svm.tanh, newdata = bank_test)
caret::confusionMatrix(bank_predict_svm.tanh, bank_test$y) # kappa = 0.17,   Sensitivity = 0.26078,    Specificity = 0.90806
```

from above 3 methods - trying adding cost param in kernel = rbfdot
#### Method 4 : kernel = rbfdot & C = 10
```{r}
bank_svm.rbf10 <- ksvm(y ~ . , data = bank_train, kernel = "rbfdot", C = 10)
bank_predict_svm.rbf10 <- predict(object = bank_svm.rbf10, newdata = bank_test)
caret::confusionMatrix(bank_predict_svm.rbf10, bank_test$y) # kappa = 0.4991,   Sensitivity = 0.47737,    Specificity = 0.96607 
```

#### Method 5 : kernel = rbfdot & C = 50
```{r}
bank_svm.rbf50 <- ksvm(y ~ . , data = bank_train, kernel = "rbfdot", C = 50)
bank_predict_svm.rbf50 <- predict(object = bank_svm.rbf50, newdata = bank_test)
caret::confusionMatrix(bank_predict_svm.rbf50, bank_test$y) # kappa = 0.4848,   Sensitivity = 0.49461,    Specificity = 0.95622
```

#### Method 6 : kernel = rbfdot & C = 100
```{r}
bank_svm.rbf100 <- ksvm(y ~ . , data = bank_train, kernel = "rbfdot", C = 100)
bank_predict_svm.rbf100 <- predict(object = bank_svm.rbf100, newdata = bank_test)
caret::confusionMatrix(bank_predict_svm.rbf100, bank_test$y) # kappa = 0.4853,   Sensitivity = 0.51509,    Specificity = 0.95034 ==> BEST
```

```{r}
confusionMatrix(bank_predict_svm.rbf100, bank_test$y, mode="prec_recall")
```

