---
title: "ExerciseLab_2"
output: html_document
---

Prerequisite
```{r}
#mtcars$cyl <- as.factor(mtcars$cyl)
#mtcars$am <- factor(mtcars$am, labels=c("Automatic","Manual"), levels=c(0,1))

```

1. Are there more automatic or manual cars?
```{r}
if (sum(mtcars$am == "Automatic") > sum(mtcars$am == "Manual")) {
  print("Automatic cars are more")
} else { 
  print("Manual cars are more")
}
```

2. Which car is heaviest, and which is fastest?
```{r}
library(plyr)
print(paste("heaviest car : ", row.names(mtcars)[which.max(mtcars$wt)]))
print(paste("fastest car : ", row.names(mtcars)[which.max(mtcars$qsec)]))
```

3. Do automatic or manual cars have on average a better mpg?
```{r}
library(dplyr)
auto.am <- filter(mtcars, am != "Manual")
manual.am <- filter(mtcars, am == "Manual")
if (mean(auto.am$mpg) > mean(manual.am$mpg)) {
  print("Automatic cars have better mpg")
} else if (mean(auto.am$mpg) < mean(manual.am$mpg)) {
  print("Manual cars have better mpg")
} else {
  print("Both type of cars have same mpg")
}
```

3. Do automatic or manual cars have on average a better mpg? (WITHOUT DPLYR)
```{r}
if (mean(mtcars$mpg[which(mtcars$am == "Automatic")]) > mean(mtcars$mpg[which(mtcars$am == "Manual")])) {
  print("Automatic cars have better mpg on average")
} else if (mean(mtcars$mpg[which(mtcars$am == "Automatic")]) < mean(mtcars$mpg[which(mtcars$am == "Manual")])) {
  print("Manual cars have better mpg on average")
} else {
  print ("Both cars have same mpg on average")
}
```

4. How many cars have above average hp?
```{r}
print(length(mtcars$hp[mtcars$hp > mean(mtcars$hp)]))

# or

print(sum(mtcars$hp > mean(mtcars$hp)))
```


5. Of the cars that have above average hp, how many have 6 cylinders?
```{r}
sum(mtcars$hp > mean(mtcars$hp) & mtcars$cyl == 6)
```


6. Of the cars that have above average hp, and 6 cylinders, how many are automatic?
Questions covering material we haven’t covered yet (in case you finish early)
```{r}
sum(mtcars$hp > mean(mtcars$hp) & mtcars$cyl == 6 & mtcars$am == "Automatic")
```


7. Make a boxplot of mpg split by no. of cylinders
```{r}
boxplot(mtcars$mpg ~ mtcars$cyl)
```


8. In the above boxplot are there outliers?
Yes, we have outlier in 8 cylinder

9. Take a random 50% sample of the dataset, and rerun questions above to see what changes
```{r}
index <- sample(1:nrow(mtcars), size = nrow(mtcars)/2)
mtcars.half <- mtcars[index,]
```

10. Split the dataset in 2, one half containing only automatics, and one half containing only manual transmissions
```{r}
mtcars.auto <- mtcars[which(mtcars$am == "Automatic"),]
mtcars.manual <- mtcars[which(mtcars$am == "Manual"),]
```

11. For each half, plot mpg against hp using the plot function
```{r}
plot(x = mtcars.auto$mpg, y = mtcars.auto$hp, main="MPG vs. HP for Automatics")
```
```{r}
plot(x = mtcars.manual$mpg, y = mtcars.manual$hp, main="MPG vs. HP for Manual")
```


12. Add an “abline” that “fits” a linear model between the two variables in the plots you just drew
```{r}
linerModel.auto <- lm(mpg ~ hp, mtcars.auto)
plot(x = mtcars.auto$hp, y = mtcars.auto$mpg, main="MPG vs. HP for Automatics")
abline(linerModel.auto)
```

```{r}
linerModel.manual <- lm(mpg ~ hp, mtcars.manual)
plot(x = mtcars.manual$hp, y = mtcars.manual$mpg, main="MPG vs. HP for Automatics")
abline(linerModel.auto)
```


Solutions will be available next week. However, I expect you to have completed the first 6, and attempted at least 1 of questions 7-12.