---
title: "Appendix 1 - King County house price prediction"
output:
  word_document: default
  html_notebook: default
---

## Part 2 - Decision Tree Regression Model

### Step 1 : Collecting Data
Same as of Part 1 : Step 1

### Step 2 : Exploring, preprocessing and cleaning the data
Primary setup
```{r setup}
knitr::opts_knit$set(root.dir = '/Users/sobil/Documents/MSC/Sem 1/Data Mining & Machine Learning/Project/Kc_house/')
remove(list = ls())
set.seed(1)
```

adding alll the libraries required
```{r}
library(RWeka)
library(rpart)
library(rpart.plot)
library(caret)
```


 Rest is same as of Part 1 : Step 2
```{r}
source("Kc_Import_Explore_Clean.R")
```

### Step 3 - Data transformation & preparation
creating training and testing dataset from exisitng sample
```{r}
indx <- createDataPartition(kc_house$price, p = 0.8, list = FALSE)

house_train <- kc_house[indx,]
house_test <- kc_house[- indx,]
```

### Step 4 : Training a model
modeling a regression tree
```{r}
house.rpart <- rpart(price ~ ., data = house_train)
```

### Step 5 : Evaluating the model
Summary of the model and description
```{r}
#house.rpart
summary(house.rpart)
```

visulazing decision tree
```{r}
rpart.plot(house.rpart, digits = 3)
```

Evaluating the model on testing data
```{r}
house.predict.rpart <- predict(house.rpart, house_test)
summary(house.predict.rpart)
summary(house_test$price)
```

checking correlation
```{r}
cor(house.predict.rpart, house_test$price)
```

cheking performace with MAE
creating a function mae to calculate Mean Absolute error
```{r}
mae <- function(actual, pred) {
  mean(abs(actual-pred))
}
```

MAE with respect to predicted values
```{r}
mae(house_test$price, house.predict.rpart)
```

MAE with respect to mean values
```{r}
mae(mean(house_train$price), house.predict.rpart)
```
RMSE - Root Mean Squared Error - log values to compare to multi-linear model 
```{r}
RMSE(log(house_test$price), log(house.predict.rpart))
```


### Step 6 : Improving the model
modeling a model tree
```{r}
house.m5p <- M5P(price ~ . , data = house_train)
#house.m5p
summary(house.m5p)
```

```{r}
house.predict.m5p <- predict(house.m5p, house_test)
summary(house.predict.m5p)
```
```{r}
cor(house.predict.m5p, house_test$price)
mae(house_test$price, house.predict.m5p)
RMSE(log(house_test$price), log(house.predict.m5p))
```
